Implicit yaddayadda

### 2011-10-13 papers/PLDI2012/implicitization.tex
exemple

\subsection{A simple, motivating example}
\label{subsec:rate_lim}
Consider the following program, adapted from \cite{Monniaux_Gonnord_SAS11}:
\lstinputlisting{rlim_int.c}
This program implements a construct commonly found in control programs (in e.g. automotive or avionics): a rate limiter. For the sake of simplicity, we chose it to be fed a nondeterministic input clamped between $[-100000,100000]$, but in a real system it would be integrated in a reactive control loop and its input connected to a complex system with unknown output range.

@@

\begin{enumerate}
	\item The first non-empty subset of paths is depicted on Figure
		\ref{fig:example-graph} Step 1. We apply Path-Focusing on this graph,
		and discover for $p_5$ the inductive invariant $\lstinline|x_old| = 0$.
		We also compute the set $P'$ of paths that have to be added at the next
		iteration. 
		For describing the different paths in the loop, we introduce the Boolean
		values $c_7, c_8, c_9$, where $c_i$ is \emph{true} when the path goes
		through the \emph{then} branch of the \emph{if-then-else} at line $i$, and
		\emph{false} when the path goes through the \emph{else} branch.
		We then add into $P'$ the following paths:
\JH{A compl\'eter}
\end{enumerate}

%Our combined technique gives us the expected invariant 
%$\lstinline|x_old| \in~[-10000,10000]$.

\begin{figure}
\label{fig:example-graph}
\centering
\begin{minipage}[c]{.10\textwidth}
\begin{tikzpicture}[->,>=stealth',auto,node distance=2.1cm,
                    semithick,font=\footnotesize]

@@

  \path [transition] 
		(n00) edge              node {} (n0);
\end{tikzpicture}
\end{minipage} 
\begin{minipage}[c]{.21\textwidth}
\begin{tikzpicture}[->,>=stealth',auto,node distance=2.1cm,
                    semithick,font=\footnotesize]

@@

  \path [transition] 
		(n00) edge  node {} (n0);
  \path [transition] 
		(n0) edge  [loop right] node [right] {$\neg c_7 \wedge c_8 \wedge \neg c_9$} (n0);
  \path [transition] 
		(n0) edge  [loop left] node [left] {$\neg c_7 \wedge c_8 \wedge \neg c_9$} (n0);
  \path [transition] 
		(n0) edge  [out=-45, in=45] node [right] {$\neg c_7 \wedge \neg c_8 \wedge c_9$} (n1);
  \path [transition] 
		(n0) edge  [out=-135, in=135] node [left] {$\neg c_7 \wedge \neg c_8 \wedge c_9$} (n1);
\end{tikzpicture}
\end{minipage}
\caption{Ascending sequence of subgraphs}
\end{figure}

>>>>

\subsection{A simple, motivating example}
\label{subsec:rate_lim}
\JH{les tests des lignes 7 et 8 ont \'et\'e remplac\'es par des in\'egalit\'es
strictes}
Consider the following program, adapted from \cite{Monniaux_Gonnord_SAS11}:
\lstinputlisting{rlim_int.c}
This program implements a construct commonly found in control programs (in e.g. automotive or avionics): a rate limiter. For the sake of simplicity, we chose it to be fed a nondeterministic input clamped between $[-100000,100000]$, but in a real system it would be integrated in a reactive control loop and its input connected to a complex system with unknown output range.

@@

\begin{enumerate}
	\item The first subset of paths is depicted on Figure
		\ref{fig:example-graph} Step 1. We apply Path-Focusing on this graph,
		and discover for $p_5$ the inductive invariant $\lstinline|x_old| = 0$.
	\item
		We then compute the set $P'$ of paths that have to be added into the
		next subgraph. 
		The image of $\lstinline|x_old| = 0$ by the path that goes from $p_5$ to
		itself, and that goes through the \emph{else} branch of each 
		\emph{if-then-else}, is $-10 \leq \lstinline|x_old| \leq 10$. This path
		is then added to our subgraph. Moreover, there is no other path whose
		image is not in $-10 \leq \lstinline|x_old| \leq 10$.
		Additionnally, the path that goes from $p_5$ to
		$p_11$, that goes through the \emph{else} branch of each 
		\emph{if-then-else} at lines 7,8 and 9, is also 
		$-10 \leq \lstinline|x_old| \leq 10$, and thus is also added into the
		new subgraph. Again, there is no new path that goes to $p_11$ with an
		image outside of $-10 \leq \lstinline|x_old| \leq 10$.
	\item
		We apply Path Focusing on this new subgraph (\ref{fig:example-graph}
		Step 2). When we have reached an inductive invariant, we narrow, and the
		result for both $p_5$ and $p_11$ is the polyhedron 
		$-10000 \leq \lstinline|x_old| \leq 10000$.
	\item Finally, we compute the new subgraph. The SMT-solver does not find any
		new path that makes the abstract values grow, and the algorithm
		terminates.
\end{enumerate}

Our technique gives us the expected invariant 
$\lstinline|x_old| \in~[-10000,10000]$. 
Here, only 3 paths over the 25 have been computed during the analysis. In
practice, depending on the order the SMT-solver returns the paths, other
feasible paths could have been added during the analysis.

\begin{figure}
\label{fig:example-graph}
\centering
\begin{minipage}[c]{.15\textwidth}
\begin{tikzpicture}[->,>=stealth',auto,node distance=2.1cm,
                    semithick,font=\footnotesize]

@@

  \path [transition] 
		(n00) edge  node {$\lstinline|x_old| \gets 0$} (n0);
\end{tikzpicture}
\end{minipage} 
\begin{minipage}[c]{.28\textwidth}
\begin{tikzpicture}[->,>=stealth',auto,node distance=2.1cm,
                    semithick,font=\footnotesize]

@@

  \path [transition] 
		(n00) edge  node {$\lstinline|x_old| \gets 0$} (n0);
  \path [transition] 
		(n0) edge  [loop left] node [left, xshift=0.2cm] {
		$\begin{array}{r}
			-10000 \leq \lstinline|x| \leq 10000 \\
			\lstinline|x_old|-10 \leq \lstinline|x| \leq \lstinline|x_old|+10 /\\
			\lstinline|x_old| \gets \lstinline|x|
		\end{array}$
		} (n0);
  \path [transition] 
		(n0) edge node [left, yshift=-0.2cm, xshift=0.2cm] {
		$\begin{array}{r}
			-10000 \leq \lstinline|x| \leq 10000 \\
			\lstinline|x_old|-10 \leq \lstinline|x| \leq \lstinline|x_old|+10 /\\
			\lstinline|x_old| \gets \lstinline|x|
		\end{array}$
		} (n1);
\end{tikzpicture}
\end{minipage}
\caption{Ascending sequence of subgraphs}
\end{figure}
